Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.Rmd
fix the different x range issue

Shape of the non-linearity pattern is another factor contributes to the non-linearity model. With different non-linearity patterns, it is expected that the visual test will respond variously.

@@

poly_conv_sim %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(RESET4_p_value < 0.05)) %>%
  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform", type == "polynomial") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(p_value < 0.05)) %>%
  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>%
  select(shape, log_effect_size, reject, conv_or_not) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject, col = shape), 
              se = FALSE, 
              method = "glm",
              method.args = list(family = binomial())) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  scale_color_manual(values = rcartocolor::carto_pal(4, "TealGrn")) +
  theme_light() +
  theme(legend.position = "bottom") +
  facet_wrap(~conv_or_not) +
  labs(col = "Non-linearity shape")
```

@@

>>>>

Shape of the non-linearity pattern is a primary factor contributes to the non-linearity model. With different non-linearity patterns, it is expected that the visual test will respond variously.

@@

dat_poly_j <- poly_conv_sim %>%
  filter(x_dist == "uniform") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(RESET4_p_value < 0.05)) %>%
  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform", type == "polynomial") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(p_value < 0.05)) %>%
  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>%
  select(shape, log_effect_size, reject, conv_or_not)

map_df(levels(dat_poly_j$shape), function(x) {

  # Adjust the range of x
  result <- data.frame(log_effect_size = seq(min(dat_poly_j$log_effect_size),
                                             max(dat_poly_j$log_effect_size), 
                                             0.01))

  # Manually fit the model
  conv_mod <- glm(reject ~ log_effect_size, 
                  family = binomial(),
                  data = filter(dat_poly_j, 
                                shape == x, 
                                conv_or_not == "Conventional"))

  # Manually fit the model
  visual_mod <- glm(reject ~ log_effect_size, 
                  family = binomial(),
                  data = filter(dat_poly_j, 
                                shape == x, 
                                conv_or_not == "Visual"))

  # Get the prediction
  result <- result %>%
    mutate(Conventional = predict(conv_mod, 
                                  newdata = result, 
                                  type = "response"),
           Visual = predict(visual_mod, 
                            newdata = result, 
                            type = "response"),
           shape = x) %>%
    pivot_longer(Conventional:Visual, names_to = "conv_or_not")

  result
}) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, value, col = shape), se = FALSE, span = 0.03) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  scale_color_manual(values = rcartocolor::carto_pal(4, "TealGrn")) +
  theme_light() +
  theme(legend.position = "bottom") +
  facet_wrap(~conv_or_not) +
  labs(col = "Non-linearity shape")
```

@@

We also investigate the impact of different heteroskedasticity patterns. In theory, the "Left-triangle" and the "Right-triangle" should have little difference from the point of view of a Breusch-Pagan test. According to Figure \ref{fig:heter-power-uniform-a}, this is indeed the case. 
Similarly, these two shapes should have the same visual test power, unless the orientation of the visual pattern is a non-negligible factor from human perspective. However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve for the "Left-triangle" shape is constantly higher than the one for the "Right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.

The "Butterfly" shape

```{r heter-power-uniform-a, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$."}

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.tex
fix the different x range issue

\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. Power curves with higher order of non-linearity are drawn with deeper colours. The default RESET tests under-perform significantly in detecting "Tripe-U" shape. To achieve the similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test. But this means the order needs to be known prior to testing.}\label{fig:poly-power-uniform-j}
\end{figure}

@@

>>>>

\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. Power curves with higher order of non-linearity are drawn with deeper colours. The default RESET tests under-perform significantly in detecting "Tripe-U" shape. To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test. But this means the order needs to be known prior to testing.}\label{fig:poly-power-uniform-j}
\end{figure}

@@

We also investigate the impact of different heteroskedasticity patterns.
In theory, the ``Left-triangle'' and the ``Right-triangle'' should have
little difference from the point of view of a Breusch-Pagan test.
According to Figure \ref{fig:heter-power-uniform-a}, this is indeed the
case. Similarly, these two shapes should have the same visual test
power, unless the orientation of the visual pattern is a non-negligible
factor from human perspective. However, it can be observed from Figure
\ref{fig:heter-power-uniform-a} that the power curve for the
``Left-triangle'' shape is constantly higher than the one for the
``Right-triangle'' shape, indicating a potential favour of orientation
by human, which is worth to be explored in future studies.

The ``Butterfly'' shape

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heter-power-uniform-a-1}

\caption{Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$.}\label{fig:heter-power-uniform-a}
\end{figure}
